Combining formal and distributional models of predicate semantics

### 2016-03-15 papers/acl2016/paper.tex
working on background
Former-commit-id: bb38d049dc5d8f97bde0d9c0e66ebe0e9a51cca9

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} introduced a
technique for learning semantic parsers with open predicate vocabularies
(hereafter referred to as K\&M).  Instead of mapping text to Freebase queries,
they applied the matrix factorization semantics used by Riedel et
al.~\shortcite{riedel-2013-mf-universal-schema} in knowledge base completion to
the task of learning predicate semantics in a semantic parser.  This method
assigns truth values to instantiated predicates like
\predicate{architect}(\entity{AndreaPalladio}) in a probabilistic database.
The models K\&M use for the predicates and entities are essentially equivalent
to modern techniques for constructing word embeddings, implicitly factorizing a
cooccurrence matrix between categories and entities, and relations and entity
pairs~\cite{levy-2014-w2v-as-mf}.  This implicit factorization has two sets of
parameters: each predicate (category or relation) has a learned $k$-dimensional
embedding $\theta$, and each entity and entity pair has a learned embedding
$\phi$.  The probability assigned to a category instance $c(e)$ and relation
instance $r(e_1, e_2)$ in the probabilistic database is given by:

@@

The probability of a predicate instance is the (sigmoided) inner product of the
corresponding predicate and entity embeddings. Thus, predicates with nearby
embeddings will have similar distributions over the entities in their
denotation.  The model is trained using a ranking objective that learns to rank
the observed entities for each query above unobserved entities.  The
probabilistic database produced by these predicate models is then input to a
reasoning system that can handle a subset of first-order logic.

In this paper, we use the system produced by K\&M, changing only the models of
predicate semantics, as described in \secref{method}.  K\&M introduced several
variations of their parser; in this work, we use their query ranking objective,
as it had the best performance, and we evaluate the model they called
``Factorization'', without ensembling it with a clustering baseline.

>>>>

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}
introduced a technique for learning semantic parsers with open
predicate vocabularies (hereafter referred to as K\&M). Instead of
mapping text directly to Freebase queries, this method parses text to
a surface logical form whose predicates are derived directly from the
words in the text. Next, a distribution over denotations for each
predicate is learned using a matrix factorization approach similar to
that of Riedel et
al.~\shortcite{riedel-2013-mf-universal-schema}. This distribution is
concisely represented using a probabilistic database, which also
enables efficient probabilistic execution of logical form queries.

This method assigns a truth probability to a predicate instance such
as \predicate{architect}(\entity{AndreaPalladio}) using probabilistic
matrix factorization. This factorization has two sets of parameters:
each category or relation has a learned $k$-dimensional embedding
$\theta$, and each entity or entity pair has a learned embedding
$\phi$. The probability assigned to a category instance $c(e)$ or
relation instance $r(e_1, e_2)$ is given by:

@@

The probability of a predicate instance is the (sigmoided) inner
product of the corresponding predicate and entity embeddings. Thus,
predicates with nearby embeddings will have similar distributions over
the entities in their denotation.

The model is trained using a query ranking objective that learns to
rank entities observed in the denotation of a logical form above
unobserved entities. The training data consists of logical form /
entity pairs $\{(\ell^i, e_^i)\}_{i=1}^n$ that are automatically
derived from an entity-linked web corpus. Training optimizes the
ranking objective function using AdaGrad, and produces the parameters
$\theta$ and $\phi$ as output. Note that this approach is similar to
modern techniques for constructing word embeddings in that it factors
a cooccurrence matrix between predicates and
entities~\cite{levy-2014-w2v-as-mf}. Given the trained predicate and
entity parameters, the system is capable of efficiently computing the
marginal probability that an entity is an element of a logical form's
denotation using approximate inference algorithms for probabilistic
databases.

This paper extends K\&M by modifying the models of predicate semantics
as described in \secref{method}. K\&M introduced several variations of
their parser; in this work, we use their query ranking objective, as
it had the best performance, and we compare against the model they
called ``Factorization,'' without ensembling it with a corpus lookup
baseline.
